---
title: "Quick Demonstration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick_Demonstration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  tidy.opts = list(width.cutoff = 60), 
  tidy = TRUE
)
 
options(rmarkdown.html_vignette.check_title = FALSE)
```

This vignette shows you some examples of using the package's main function, [`create_pal`](https://genchanghsu.github.io/PalCreatoR/reference/create_pal.html). You will learn how to make good use of its functionality to create your own awesome palette.   

We will play around with the [`create_pal`](https://genchanghsu.github.io/PalCreatoR/reference/create_pal.html) function using the image below (photo credit: Gen-Chang Hsu). This stonecrop species (*Sedum morrisonense* Hayata) is endemic to Taiwan and commonly seen at mid- and high-altitudes. It thrives in the rock cracks and crevices, growing low and spreading out to avoid strong wind. It is also characterized by its densely-clumped, fingernail-like succulent leaves, which are used for water storage in addition to photosynthesis. During summertime, blossoms of bright yellow flowers will burst into bloom.

<img src = "./inst/Alpine Flower.JPG" align = "center" height = "400" />

<br>

# Part 1. Create the Palette
Now let's start with a simple example:
```{r, fig.show = "hold", out.width = "50%", fig.width = 3.5, fig.height = 4}
# Get the path of the image
library(PalCreatoR) 
image_path <- system.file("Alpine Flower.JPG", package = "PalCreatoR")

# Create a basic palette with 5 colors
My_pal_5 <- create_pal(image = image_path, n = 5, title = "My Palette (n = 5)")

print(My_pal_5)
```

<br>

More colors:
```{r, fig.show = "hold", out.width = "50%", fig.width = 3.5, fig.height = 4}
My_pal_10 <- create_pal(image = image_path, n = 10, title = "My Palette (n = 10)")

print(My_pal_10)
```

<br>

By default, the colors in the palette are not checked for colorblind safe. To get the colorblind-friendly colors, set `colorblind = TRUE`:
```{r, fig.show = "hold", out.width = "50%", fig.width = 3.5, fig.height = 4}
My_pal_clb <- create_pal(image = image_path, n = 5, colorblind = TRUE, title = "My Palette \n (colorblind-friendly)")

print(My_pal_clb)
```

<br>

The colors in the palette are not ordered by default. You can sort the colors in the [HSV color space](https://en.wikipedia.org/wiki/HSL_and_HSV). Three options are available:
<ol>
  <li>`sort = "hue"`: the colors are sorted by hue ^[Hue represents the angle of the color (in degree) on the RGB color circle. A hue value of 0° gives red, 120° green, and 240° blue.] in an ascending order (i.e., from red, yellow, green, to cyan, blue, and magenta);</li>
  <li>`sort = "saturation"`: the colors are sorted by saturation ^[Saturation controls the amount of color used. A saturation value of 100% will be the purest color possible for the given hue; a saturation value of 0% results in grayscale color.] in a descending order (i.e., from pure to gray);</li>
  <li>`sort = "value"`: the colors are sorted by value ^[Value indicates the brightness of the color. A value of 100% has no black mixed into the color; a value of 0% gives pure black.] in a descending order (i.e., from bright to dark).</li>
</ol>

Let's visualize the effects of color sorting:
```{r, fig.show = "hold", out.width = "50%", fig.width = 3.5, fig.height = 4}
# Unsorted
My_pal_none <- create_pal(image = image_path, n = 5, sort = "none", title = "My Palette \n (unsorted)")

# By hue
My_pal_hue <- create_pal(image = image_path, n = 5, sort = "hue", title = "My Palette \n (sorted by hue)")

# By saturation
My_pal_saturation <- create_pal(image = image_path, n = 5, sort = "saturation", title = "My Palette \n (sorted by saturation)")

# By value
My_pal_value <- create_pal(image = image_path, n = 5, sort = "value", title = "My Palette \n (sorted by value)")
```

<br>

Another novel feature in the [`create_pal`](https://genchanghsu.github.io/PalCreatoR/reference/create_pal.html) function is that you can specify `method = "Gaussian_mix"` to apply multivariate Gaussian mixture modeling (GMM)
instead of the default kmeans algorithm to extract the representative colors from the image: 

```{r, fig.show = "hold", out.width = "50%", fig.width = 3.5, fig.height = 4}
# kmeans
My_pal_kmeans <- create_pal(image = image_path, n = 5, method = "kmeans", sort = "value", title = "My Palette \n (kmeans)")

# Gaussian mixture modeling
My_pal_Gaussian_mix <- create_pal(image = image_path, n = 5, method = "Gaussian_mix", sort = "value", title = "My Palette \n (GMM)")
```

Not much difference in this case.

<br>

Finally, if you do not want to see your palette, set `show.pal = FALSE`. This will return only the color code vector. 









# Part 2. Modify the Palette
You can modify the alpha transparency of the colors in the palette using the helper function [`modify_pal`](https://genchanghsu.github.io/PalCreatoR/reference/modify_pal.html). Simply pass the palette and a vector of alpha values into the function, and you will get the modified colors!

```{r, fig.show = "hold", out.width = "50%", fig.width = 3.5, fig.height = 4}
# kmeans
My_pal_kmeans <- create_pal(image = image_path, n = 5, method = "kmeans", sort = "value", title = "My Palette \n (kmeans)")

# Gaussian mixture modeling
My_pal_Gaussian_mix <- create_pal(image = image_path, n = 5, method = "Gaussian_mix", sort = "value", title = "My Palette \n (GMM)")
```

In general, the number of palette and the number of alpha values should be the same. Any extra elements in the longer vector will be omitted to match the elements in the shorter one. However, there are two exceptions: <ol>
  <li>A single color with multiple alpha values: If you have a single color with multiple alpha values, then you will get a sequence of that color with different transparencies. This is useful for generating a color gradient.</li>
```{r, fig.show = "hold", out.width = "50%", fig.width = 3.5, fig.height = 4}
My_pal_10 <- create_pal(image = image_path, n = 10, title = "My Palette (n = 10)")

print(My_pal_10)
```

  <li>Multiple colors with a single alpha values: If you have multiple colors with a single alpha value, then you will get the same set of colors adjusted by the same degree of transparency. This is useful for toning down the colors in a uniform manner.</li>
```{r, fig.show = "hold", out.width = "50%", fig.width = 3.5, fig.height = 4}
My_pal_10 <- create_pal(image = image_path, n = 10, title = "My Palette (n = 10)")

print(My_pal_10)
```

 






<br>

# Part 3. Use the Palette
Finally, let's use the palettes in the real figures:
```{r, tidy = FALSE}
library(ggplot2)
library(ggpubr)

# Create the palettes
My_pal1 <- create_pal(image = image_path, n = 10, sort = "value", colorblind = FALSE, show.pal = F)
My_pal2 <- create_pal(image = image_path, n = 10, sort = "value", colorblind = TRUE, show.pal = F)

# The golden spiral
spiral_df <- data.frame(x = seq(0, 13, length.out = 5000)) %>%
  mutate(y = exp(0.30635*x)) %>%
  mutate(x = rev(x),
         y = rev(y),
         polar_x = y*cos(x),
         polar_y = y*sin(x),
         x_start = rep(0, 5000),
         y_start = rep(0, 5000))

# Plot the spirals (package logo)
spiral_p1 <- ggplot(spiral_df) +
  geom_segment(aes(x = x_start, y = y_start, xend = polar_x, yend = polar_y),
               color = rep(My_pal1, each = 500),
               size = 1) +
  geom_point(aes(x = polar_x, y = polar_y),
             color = "black",
             size = 0.1) +
  coord_fixed(ratio = 1) +
  theme_void()

spiral_p2 <- ggplot(spiral_df) +
  geom_segment(aes(x = x_start, y = y_start, xend = polar_x, yend = polar_y),
               color = rep(My_pal2, each = 500),
               size = 1) +
  geom_point(aes(x = polar_x, y = polar_y),
             color = "black",
             size = 0.1) +
  coord_fixed(ratio = 1) +
  theme_void()

# Arrange the spirals
ggarrange(spiral_p1, spiral_p2)
# Dazzling effect
spiral_p1_dazzle <- ggplot(spiral_df) +
  geom_segment(aes(x = x_start, y = y_start, xend = polar_x, yend = polar_y),
               color = rep(My_pal1, 500),
               size = 1) +
  geom_point(aes(x = polar_x, y = polar_y),
             color = "black",
             size = 0.1) +
  coord_fixed(ratio = 1) +
  theme_void()

spiral_p2_dazzle <- ggplot(spiral_df) +
  geom_segment(aes(x = x_start, y = y_start, xend = polar_x, yend = polar_y),
               color = rep(My_pal2, 500),
               size = 1) +
  geom_point(aes(x = polar_x, y = polar_y),
             color = "black",
             size = 0.1) +
  coord_fixed(ratio = 1) +
  theme_void()

# Arrange the spirals
ggarrange(spiral_p1_dazzle, spiral_p2_dazzle)
# Continuous color spectrum
My_pal3 <- create_pal(image = image_path, n = 5000, sort = "value", colorblind = FALSE, show.pal = F) 
# This might take a while to run

spiral_p3 <- ggplot(spiral_df) +
  geom_segment(aes(x = x_start, y = y_start, xend = polar_x, yend = polar_y),
               color = rep(My_pal3),
               size = 1) +
  geom_point(aes(x = polar_x, y = polar_y),
             color = "black",
             size = 0.1) +
  coord_fixed(ratio = 1) +
  theme_void()

spiral_p3
# Color gradients with different alpha transparencies
My_pal4 <- create_pal(image = image_path, n = 5, sort = "value", colorblind = FALSE, show.pal = F) 

My_new_pal <- modify_pal(pal = My_pal4[1], alpha = seq(0.1, 1.0, 0.1), show.pal = T)

spiral_new_pal <- ggplot(spiral_df) +
  geom_segment(aes(x = x_start, y = y_start, xend = polar_x, yend = polar_y),
               color = rep(My_new_pal, each = 1000),
               size = 1) +
  geom_point(aes(x = polar_x, y = polar_y),
             color = "black",
             size = 0.1) +
  coord_fixed(ratio = 1) +
  theme_void()

spiral_new_pal

```

<br>

Interested? Grab your own image and try it out now!

Have fun and enjoy!!!




